---
title: "geom_hist: Histogram with Points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geom_hist: Histogram with Points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

The `geom_hist()` function is a variant of `geom_histogram()` that displays histogram bin counts using points instead of bars. This can be useful when you want to emphasize the bin counts rather than the bar heights, or when you want to overlay multiple histograms without bar overlap.

## Basic Usage

The simplest use of `geom_hist()` requires only an x aesthetic:

```{r, message=FALSE, warning=FALSE}
library(mfptools)
library(ggplot2)

ggplot(mtcars, aes(mpg)) +
  geom_hist(bins = 20)
```

## Log Scale for Large Dynamic Range

When bin counts span several orders of magnitude, using a log scale on the y-axis can make the differences more visible:

```{r, message=FALSE, warning=FALSE}
# Create a dataset with large dynamic range
set.seed(42)
data_large <- data.frame(
  value = c(
    rnorm(10000, mean = 50),
    rnorm(1000, mean = 100),
    rnorm(100, mean = 150)
  )
)

ggplot(data_large, aes(value)) +
  geom_hist(bins = 30) +
  scale_y_log10()
```

## Customizing the Y Aesthetic

You can also use other computed variables like `density`:

```{r, message=FALSE, warning=FALSE}
ggplot(mtcars, aes(mpg)) +
  geom_hist(aes(y = after_stat(density)), bins = 20)
```

## Dropping Zero Counts

By default, `geom_hist()` drops bins with zero counts (`drop.zeros = TRUE`). You can include them by setting `drop.zeros = FALSE`:

```{r, message=FALSE, warning=FALSE}
# Create data with gaps
data_gaps <- data.frame(
  x = c(1:5, 10:15)
)

ggplot(data_gaps, aes(x)) +
  geom_hist(bins = 10, drop.zeros = FALSE)
```

## Position Adjustment

Like other geoms, you can use position adjustments to overlay multiple histograms:

```{r, message=FALSE, warning=FALSE}
ggplot(mtcars, aes(mpg, colour = factor(cyl))) +
  geom_hist(bins = 20, position = "identity", alpha = 0.5)
```
